# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Load the activity.csv to a dataframe.
```{r echo=TRUE}
activity <- read.csv(paste(getwd(), "/activity.csv", sep = ""))
```

## What is mean total number of steps taken per day?
Get the total steps for each day and put that in its own dataframe.
```{r echo=TRUE}
totals <- aggregate(activity[complete.cases(activity$steps),]$steps, list(period=activity[complete.cases(activity$steps),]$date), sum)
```

Make a histogram from to show the steps taken per day.
```{r echo=TRUE}
hist(totals$x, main="Steps Taken Per Day", xlab="Steps Taken Per Day")
```

Print the mean and median number of steps taken per day.
```{r echo=TRUE}
mean(totals$x)
median(totals$x)
```

## What is the average daily activity pattern?
Build a dataframe containing the mean number of steps per interval.
```{r echo=TRUE}
interval <- aggregate(activity[complete.cases(activity$steps),]$steps, list(interval=activity[complete.cases(activity$steps),]$interval), mean)
```

Plot the number of steps per five minute interval.
```{r echo=TRUE}
plot(interval$interval, interval$x,  type="l", pch="", ylab="Mean Steps Taken",
     xlab = "Five Minute Interval")
```

Find the time interval where the most steps are taken.
```{r echo=TRUE}
interval[interval$x == max(interval$x),]$interval
```

Apparently, most people are walking between 8:35AM - 8:40AM

## Imputing missing values
Merge the original dataset with the mean number of steps per interval.
```{r echo=TRUE}
full <- merge(activity,interval,by="interval")
```

If the number of steps is NA, replace it with the mean number of steps for that interval.
```{r echo=TRUE}
fixed <- transform(full,steps=ifelse(is.na(steps),x,steps))
```

Follow the same steps to produce the histogram except with the new fill dataframe.
```{r echo=TRUE}
totals2 <- aggregate(fixed[complete.cases(fixed$steps),]$steps, list(period=fixed[complete.cases(fixed$steps),]$date), sum)
hist(totals2$x, main="Steps Taken Per Day", xlab="Steps Taken Per Day")
mean(totals2$x)
median(totals2$x)
```

You can see the median went up slightly and the histogram maintained a similar shape (although the frequencies are up).

## Are there differences in activity patterns between weekdays and weekends?
Add a new column to the dataframe that determines if the record is during a weekday or weekend
```{r echo=TRUE}
fixed <- transform(fixed,weekday=ifelse(weekdays(as.Date(date))=="Sunday"|weekdays(as.Date(date))=="Saturday","Weekend","Weekday"))
```

Build a dataframe containing the mean number of steps per interval per day type.
```{r echo=TRUE}
interval2 <- aggregate(fixed[complete.cases(fixed$steps),]$steps, list(interval=fixed[complete.cases(fixed$steps),]$interval, day=fixed[complete.cases(fixed$steps),]$weekday), mean)
```

Plot the mean steps per interval during the weekdays compared to the same during the weekends
```{r echo=TRUE}
par(mfrow=c(2,1))
plot(interval2[interval2$day=="Weekday",]$interval, interval2[interval2$day=="Weekday",]$x,  type="l", pch="", ylab="Mean Steps Taken",
     xlab = "Five Minute Interval")
title(main="Weekday", col.main="blue", font.main=2)
plot(interval2[interval2$day=="Weekend",]$interval, interval2[interval2$day=="Weekend",]$x,  type="l", pch="", ylab="Mean Steps Taken",
     xlab = "Five Minute Interval")
title(main="Weekend", col.main="blue", font.main=2)
```

It can be seen that there is more walking happening during the middle hours of the weekdays.